Portfolios of Subgraph Isomorphism Algorithms

### 2015-12-17 paper/paper.tex
Minor editing

In Section~\ref{sec:algs}, we experimentally compare eight state-of-the-art
algorithms. We introduce a large benchmark set composed of 5725 instances
grouped into twelve classes. Ten of these classes have been introduced in the
experimental study reported by McCreesh and Prosser~\cite{McCreesh:2015}. We
evaluate the algorithms on this benchmark set, and show that they have very
complementary performance. In particular, we show that depending on the CPU
time limit, different algorithms achieve the best performance on the entire
benchmark set.

In Section~\ref{sec:algsel}, we describe the features that are used to describe
instances, and we describe our algorithm selection approach. It combines a
presolving step, which allows us to very quickly solve easy instances, with the
algorithm selection approach \LLAMA~\cite{kotthoff_llama_2013}.

In Section~\ref{sec:algsel-exps}, we experimentally evaluate our selection
approach, and show that it is able to close more than 50\% of the gap between
the single best and the virtual best solver.

We conclude and give directions for future work in Section~\ref{sec:concs}.

@@

Subgraph isomorphism problems may be solved by a systematic exploration of the search space composed
of all possible injective matchings from $N_p$ to $N_t$: starting from an empty matching, one
incrementally extends a partial matching by matching a non-matched pattern node to a non-matched
target node until either some edges are not matched by the current matching (so the search must
backtrack to a previous choice point and go on with another extension), or all pattern nodes have
been matched (a solution has been found). To reduce the search space, this exhaustive exploration is
combined with filtering techniques that aim at removing candidate pairs of non-matched
pattern-target nodes $(u,v)\in N_p\times N_t$. Different filtering techniques may be considered;
some are stronger than others (they remove more candidate pairs), but also have higher time
complexities.

@@

Various filtering techniques have been tried for edge constraints. Ullman~\cite{ullman} introduced a
filtering which ensures that for each pattern edge $(u,u')\in E_p$ and each candidate pair $(u,v)\in
N_p\times N_t$, there exists a candidate couple $(u',v')\in N_p\times N_t$ such that $(v,v')$ is a
target edge. Candidate couples $(u,v)$ that do not satisfy this property are iteratively removed
until a fixed point is reached. This filtering (called \emph{Arc Consistency}) removes more
candidate pairs than Forward-Checking, but it is also more time consuming as it
runs in ${\cal
O}(e_p\cdot n_t^2)$ when using AC4~\cite{MH86}.

@@

\begin{itemize}[nosep]
\item by Battiti and Mascia~\cite{battiti-mascia07}, to remove candidate pairs $(u,v)\in N_p\times
    N_t$ such that the number of paths starting from  pattern node $u$ is greater than the number of
    paths starting from  target node $v$;
\item by Audemard et al.~\cite{Audemard:2014} to generalise the locally all-different constraint
    proposed by Solnon~\cite{Solnon:2010} so that it ensures that a subset of pattern nodes can be
    matched with all different compatible target nodes, where compatibility is defined with respect
    to invariant properties;
\item by McCreesh and Prosser~\cite{McCreesh:2015} to filter the set of candidate couples before
    starting the search, and to generate additional implied adjacency-like constraints which are
    processed during search.
\end{itemize}

@@

\subsection{Novel Algorithms}

@@

Table~\ref{tab:res} shows the performance of our algorithm selection approach, compared to two
baselines, on the reduced set of 2336 instances. The virtual best solver is the oracle predictor
that, for each instance, chooses the best solver from our portfolio. This is the upper bound of what
an algorithm selection approach can achieve. The single best solver is the one solver from the
portfolio that has the overall best performance across the entire set of instances, at the CPU time
limit of \SI{e8}{\ms}, i.e.\ \GlasgowTwo. We consider it a lower bound on the performance of the
algorithm selection approach.  We are able to solve 28 more instances than the single best solver
within the timeout, with only an additional 18 to the virtual best. In terms of average performance,
we are able to more than \SI{50}{\percent} of the gap between the single best and the virtual
best solver.

Figure~\ref{fig:portfolio-ecdf} shows the number of solved instances over time for the individual
solvers, the virtual best solver, and the \LLAMA algorithm selection approach. The performance of
\VFtwo is far behind all other solvers.  The algorithm selection model does not perform well for
instances that can be solved quickly because of the overhead incurred through feature computation.
As the instances become more difficult to solve, its performance improves.

>>>>

In Section~\ref{sec:exps}, we experimentally compare eight state-of-the-art
algorithms. We introduce a large benchmark set composed of 5725 instances
grouped into twelve classes. Ten of these classes were considered in the
experimental study reported by McCreesh and Prosser~\cite{McCreesh:2015}; two
are new. We evaluate the algorithms on this benchmark set, and show that they
have very complementary performance. In particular, we show that depending on
the CPU time limit, different algorithms achieve the best performance on the
entire benchmark set.

In Section~\ref{sec:algsel}, we discuss the features that are used to describe
instances, and we describe our algorithm selection approach. It combines a
presolving step, which allows us to easy instances very quickly, with the
algorithm selection approach \LLAMA~\cite{kotthoff_llama_2013}.
In Section~\ref{sec:algsel-exps}, we experimentally evaluate our selection
approach, and show that it is able to close more than 50\% of the gap between the
single best and the virtual best solver.
We conclude and give directions for future work in Section~\ref{sec:concs}.

@@

Subgraph isomorphism problems may be solved by a systematic exploration of the search space consisting
of all possible injective matchings from $N_p$ to $N_t$: starting from an empty matching, one
incrementally extends a partial matching by matching a non-matched pattern node to a non-matched
target node until either some edges are not matched by the current matching (so the search must
backtrack to a previous choice point and go on with another extension), or all pattern nodes have
been matched (a solution has been found). To reduce the search space, this exhaustive exploration is
combined with filtering techniques that aim at removing candidate pairs of non-matched
pattern-target nodes $(u,v)\in N_p\times N_t$. Different filtering techniques may be considered;
some are stronger than others (they remove more candidate pairs), but also have higher time
complexities.

@@

Various filtering techniques have been tried for edge constraints. Ullman~\cite{ullman} introduced a
filtering which ensures that for each pattern edge $(u,u')\in E_p$ and each candidate pair $(u,v)\in
N_p\times N_t$, there exists a candidate pair $(u',v')\in N_p\times N_t$ such that $(v,v')$ is a
target edge. Candidate pairs $(u,v)$ that do not satisfy this property are iteratively removed
until a fixed point is reached. This filtering (called \emph{Arc Consistency}) removes more
candidate pairs than Forward-Checking, but it is also more time consuming as it
runs in ${\cal
O}(e_p\cdot n_t^2)$ when using AC4~\cite{MH86}.

@@

\begin{itemize}[nosep]
\item by Battiti and Mascia~\cite{battiti-mascia07}, to remove candidate pairs $(u,v)\in N_p\times
    N_t$ such that the number of paths starting from  pattern node $u$ is greater than the number of
    paths starting from  target node $v$;
\item by Audemard et al.~\cite{Audemard:2014} to generalise the locally all-different constraint
    proposed by Solnon~\cite{Solnon:2010} so that it ensures that a subset of pattern nodes can be
    matched with all different compatible target nodes, where compatibility is defined with respect
    to invariant properties;
\item by McCreesh and Prosser~\cite{McCreesh:2015} to filter the set of candidate pairs before
    starting the search, and to generate additional implied adjacency-like constraints which are
    processed during search.
\end{itemize}

@@

\subsection{New Algorithms}

@@

Table~\ref{tab:res} shows the performance of our algorithm selection approach, compared to two
baselines, on the reduced set of 2336 instances. The virtual best solver is the oracle predictor
that, for each instance, chooses the best solver from our portfolio. This is the upper bound of what
an algorithm selection approach can achieve. The single best solver is the one solver from the
portfolio that has the overall best performance across the entire set of instances, at the CPU time
limit of \SI{e8}{\ms}, i.e.\ \GlasgowTwo. We consider it a lower bound on the performance of the
algorithm selection approach.  We are able to solve 28 more instances than the single best solver
within the timeout, with only an additional 18 to the virtual best. In terms of average performance,
we are able to close more than \SI{50}{\percent} of the gap between the single best and the virtual
best solver.

Figure~\ref{fig:portfolio-ecdf} shows the number of solved instances over time
for the individual solvers, the virtual best solver, and the \LLAMA algorithm
selection approach. The algorithm selection model does not perform well for
instances that can be solved quickly because of the overhead incurred through
feature computation.  As the instances become more difficult to solve, its
performance improves.
